---
title: "Results"
format: pdf
editor: visual
bibliography: references.bib
---

```{r}
#| echo: false

#load packages
pacman::p_load(carData, 
               dplyr, 
               ggplot2, 
               knitr, 
               papaja)
```

## Results

In this section, we will present our analysis results using data from a study on volunteering [@cowles1987]. This data is available in the [@carData] package.

### Descriptive statistics and plots

Number of Volunteers by binary sex variable:

```{r}
#| tbl-cap: Volunteering Counts by Sex
#| tbl-cap-location: bottom
#| label: tbl-volbysex

Cowles %>% 
  group_by(sex, volunteer) %>% 
  tally() %>% 
  group_by(sex) %>% 
  mutate(Percent = n/sum(n)*100) %>% 
  kable(digits=2)

```

Average and SD of personality scores by volunteering.

```{r}
#| tbl-cap: Average personality scores by volunteering
#| tbl-cap-location: bottom
#| label: tbl-scoresbyvol

Cowles %>% 
  group_by(volunteer) %>% 
  summarize(Neuroticism_Avg = mean(neuroticism), 
            Neuroticism_SD = sd(neuroticism),
            Extraversion_Avg = mean(extraversion), 
            Extraversion_SD = sd(extraversion)) %>% 
  kable(digits=2)
```

Look at the distribution of data

```{r}
#|label: fig-extraversionbysexvol
#|fig-cap: Distribution of extraversion scores by volunteer status and sex
#|fig-cap-location: bottom

Cowles %>% 
  ggplot(aes(x=extraversion, fill=volunteer)) +
  geom_density(alpha=.5) +
  facet_wrap(~sex)+
  theme_classic()

```

```{r}
#|label: fig-neuroticismbysexvol
#|fig-cap: Distribution of neuroticism scores by volunteer status and sex
#|fig-cap-location: bottom

Cowles %>% 
  ggplot(aes(x=neuroticism, fill=volunteer)) +
  geom_density(alpha=.5) +
  facet_wrap(~sex)+
  theme_classic()

```

### Do personality scores differ for those who do and do not volunteer?

Independent t-tests were used to determine whether personality scores differed by whether participants volunteered or not.

```{r}
extrat <- apa_print(t.test(extraversion ~ volunteer, data=Cowles))
neurot <- apa_print(t.test(neuroticism ~ volunteer, data=Cowles))
```

Extraversion differed between those who volunteered (`r extrat$full_result`). However, scores on neuroticism did not differ between groups (`r neurot$full_result`).

### What is the relationship between extraversion and neuroticism in this sample?

Do these scores correlate in this sample in the same way for volunteers and non-volunteers?

```{r}
#| label: fig-extneurscatter
#| fig-cap: Relationship between extraversion and neuroticism in the sample. 
#| fig-cap-location: bottom
#| message: false

Cowles %>% 
  ggplot(aes(x=extraversion, y=neuroticism)) +
  geom_point(aes(color=volunteer), position=position_jitter(width = .5, height=.5)) +
  geom_smooth(method="lm", aes(color=volunteer), se=FALSE) +
  theme_classic()
```

```{r}
#| results: asis

lmtest <- apa_print(lm(neuroticism ~ extraversion * volunteer, data = Cowles))

lmtest$table %>% 
  kable(col.names = c("Predictor", "$b$", "95\\% CI", "$t$", "$\\mathit{df}$", "$p$"))
```

## References
